The \emph{FrobeniusThresholds} package for \emph{Macaulay2}

### 2019-06-20 Fsing/FrobeniusThresholdsPaper/FrobeniusThresholds.tex
Changes to paper

Consider a nonzero polynomial $f$ and a natural number $e$.  If $f$ does not vanish at the origin, then set $\nu_f^{\idealm}(p^e) := \infty$.  Otherwise, $f \in \idealm$, and we instead define
\[ \nu_f^{\idealm}(p^e) \coloneqq  \max \{ n>0 : f^n \notin \idealm^{[p^e]} \}, \]
where $\idealm^{[p^e]}$ denotes the $p^e$-th Frobenius power of $\idealm$, that is, the ideal generated by the $p^e$-th powers of the elements of $\idealm$.
%It is not hard to see that this number is the largest power of $f$ that is part of a free basis for the localization of the ambient polynomial ring at the origin, over its subring of $p^e$-th powers.
%As noted earlier, this number also has a natural interpretation in terms of $F$-purity of pairs.
Following our outline, we define
\[ \ft^{\idealm}(f) \coloneqq  \lim_{e \to \infty} \frac{ \nu_f^{\idealm}(p^e)}{p^e}. \]

@@

We conclude this subsection by briefly reviewing some natural generalizations.  Suppose that $I$ and $J$ are ideals.  If $I$ is contained in the radical of $J$, then we set
%
\[ \nu_I^J(p^e) \coloneqq \max \{ n : I^n \not\subseteq J^{[p^e]} \}. \]
Otherwise, we set $\nu_I^J(p^e) \coloneqq \infty$.
This clearly generalizes the quantity $\nu_f^{\idealm}(p^e)$ considered earlier, and we call
\[ \ft^J(I) \coloneqq  \lim_{e \to \infty} \frac{ \nu_I^J(p^e)}{p^e} \]
the \emph{$F$-threshold of $I$ with respect to $J$}. 
The value $\ft^{\idealm}(I)$ is called the \emph{$F$-pure threshold of $I$} at the origin, and if $I = \langle f \rangle$ is principal,  $\ft^{J}(f) \coloneqq  \ft^J(I)$ is called the \emph{$F$-threshold of $f$ with respect to $J$}.
Like $F$-pure thresholds, $F$-thresholds are rational (when finite), and can be characterized in terms of test ideals.

@@

\medskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i1 : R = ZZ/11[x,y];

@@

i4 : nu(1, I, J)

o4 = 24

@@

i6 : nu(1, f, J)

o6 = 10
\end{MyVerbatim}
}

@@

We now describe the options available for the function \texttt{nu}.
As pointed out in the introduction, if $f \in \mathfrak{m}$, the values $\nu^{\mathfrak{m}}_f(p^e)$ can be recovered from the $F$-pure threshold of $f$.
This is used to speed up computations for certain polynomials whose $F$-pure thresholds can be computed quickly via specialized algorithms or formulas, namely diagonal polynomials, binomials, forms in two variables, and products of factors in simple normal crossing (see Section~\ref{sec.FPT}).
This feature can be disabled by setting the option \texttt{UseSpecialAlgorithms} (default value \texttt{true}) to \texttt{false}; in Section~\ref{ss: UseSpecialAlgorithms} we discuss situations in which this may be desirable.

\medskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i13 : R = ZZ/17[x,y,z];

i14 : f = x^3 + y^4 + z^5; -- a diagonal polynomial

i15 : time nu(5, f)
     -- used 0.00573794 seconds

o15 = 1085772

i16 : time nu(5, f, UseSpecialAlgorithms => false)
     -- used 0.492965 seconds

o16 = 1085772
\end{MyVerbatim}
}
\medskip

@@

\medskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i17 : R = ZZ/5[x,y,z];

i21 : m = ideal(x, y, z);

i22 : time nu(2, m, m^2) -- uses binary search (default)
     -- used 1.71206 seconds

o22 = 97

i23 : time nu(2, m, m^2, Search => Linear) -- but linear seach gets luckier
     -- used 0.536135 seconds

o23 = 97
\end{MyVerbatim}
}
\medskip

@@

\medskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i24 : nu(5, f, ReturnList => true)

o24 = {0, 1, 8, 44, 224, 1124}

o24 : List
\end{MyVerbatim}
}
\medskip

@@

\medskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i26 : R = ZZ/11[x,y,z];

i27 : f = x^3 + y^3 + z^3 + x*y*z;

i28 : time nu(3, f) -- ContainmentTest is set to FrobeniusRoot, by default
     -- used 0.127375 seconds

o28 = 1209

i29 : time nu(3, f, ContainmentTest => StandardPower)
     -- used 9.57149 seconds

o29 = 1209
\end{MyVerbatim}
}
\medskip

@@

\medskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i30 : R = ZZ/3[x,y];

i31 : m = ideal(x, y);

o31 : Ideal of R

i32 : nu(3, m^5)

o32 = 10

i33 : nu(3, m^5, ContainmentTest => FrobeniusPower)

o33 = 8
\end{MyVerbatim}
}
\medskip

@@

\medskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i34 : R = ZZ/7[x,y];

i35 : f = (x - 1)^3 - (y - 2)^2;

i36 : nu(1, f)

o36 = infinity

o36 : InfiniteNumber

i37 : nu(1, f, AtOrigin => false)

o37 = 5
\end{MyVerbatim}
}

@@

The \emph{FrobeniusThresholds} package contains methods to test candidate values for $F$-pure thresholds and $F$-jumping numbers, even in some singular rings.
Consider a  $\mathbb{Q}$-Gorenstein ring $R$ of characteristic $p>0$, whose index is not divisible by $p$.
Given a parameter $t\in \mathbb{Q}$ and an element $f$ of $R$, the command \texttt{isFPT(t, f)} checks whether $t$ is the $F$-pure threshold of $f$, while \texttt{compareFPT(t, f)} provides further information, returning {\tt-1}, \texttt{0}, or \texttt{1} when $t$ is less than, equal to, or greater than the $F$-pure threshold of $f$.  Setting the option \texttt{AtOrigin} to \texttt{true} tells the function to consider the $F$-pure threshold at the origin.  % The following example illustrates this functionality.

\medskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i1 : R = ZZ/11[x, y, z]/(x^2 - y*(z - 1));

@@

o4 = false
\end{MyVerbatim}
}
\medskip

% The general method applied calls upon functionality from the \emph{TestIdeals} package.
% The functions \texttt{testIdeal} and \texttt{FPureModule} therein are used to compute the test ideals of $f$ with respect to the parameter $t$, and that with respect to $t - \varepsilon$, for $0<\varepsilon\ll 1$. For instance, as noted in the introduction, when $R$ is a polynomial ring, $t$ is the $F$-pure threshold of $f$ if and only if $\tau(R, f^t)$ is contained in the homogeneous maximal ideal, but
% $\tau(R, f^{t-\varepsilon})$ is not.
% Note that for $R$ strongly $F$-regular and Gorenstein, \texttt{FPureModule(t, f)} precisely computes $\tau(R, f^{t-\varepsilon})$.

The general method applied calls upon functionality from the \emph{TestIdeals} package.
The functions \texttt{testIdeal} and \texttt{FPureModule} therein are used to compute the test ideals of $f$ with respect to the parameter $t$, and that with respect to $t - \varepsilon$, for $0<\varepsilon\ll 1$.
By comparing these test ideals, we can decide whether $t$ is the $F$-pure threshold of $f$.
For instance, as noted in the introduction, when $R$ is a polynomial ring, $t$ is the $F$-pure threshold of $f$ at the origin if and only if $\tau(R, f^t)$ is contained in the homogeneous maximal ideal, but $\tau(R, f^{t-\varepsilon})$ is not.
%Note that for $R$ strongly $F$-regular and Gorenstein, \texttt{FPureModule(t, f)} precisely computes $\tau(R, f^{t-\varepsilon})$.

Since not only the $F$-pure thresholds, but also the higher $F$-jumping numbers, are determined by containment conditions on test ideals, the functionality is extended to determine whether a given number is an $F$-jumping number.
Analogous syntax is used, as shown in the following example.

\medskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i5 : R = ZZ/13[x, y];

@@

o7 = true
\end{MyVerbatim}
}

@@

\medskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i1 : R = ZZ/5[x,y,z];

@@

The \texttt{fpt} function has an option called \texttt{UseSpecialAlgorithms}, which, when set to \texttt{true} (its default value), tells \texttt{fpt} to first check whether $f$ is a diagonal polynomial, a binomial, a form in two variables, or a product of factors in simple normal crossing, in that order.

When $f$ is a diagonal polynomial, a binomial, or a form in two variables, algorithms of Hern\'andez \cite{HernandezFInvariantsOfDiagonalHyp, HernandezFPureThresholdOfBinomial}, or Hern\'andez and Teixeira \cite{HernandezTeixeiraFThresholdFunctions},  are executed to compute the $F$-pure threshold.
%Algorithms for computing $F$-pure thresholds of binomials also appear in \cite{ShibutaTakagiLCThresholds}, though they are only valid in certain special characteristics.

\medskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i4 : fpt(x^17 + y^20 + z^24) -- a diagonal polynomial

@@

o7 : QQ
\end{MyVerbatim}
}
\medskip

The algorithm for computing the $F$-pure threshold of a binary form $f$ requires factoring $f$ into linear forms, and that may be difficult or impossible when that factorization occurs in a Galois field of excessively large order.
This is a situation when the user will want to set the option \texttt{UseSpecialAlgorithms} to \texttt{false}.
However, when a factorization is already known, instead of passing $f$ to \texttt{fpt}, the user can pass a list containing all the pairwise coprime linear factors of $f$ to \texttt{fpt}, and a list containing their respective multiplicities.

\medskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i8 : fpt({x, y, x + y, x + 3*y}, {2, 6, 9, 10}) == oo

o8 = true
\end{MyVerbatim}
}
\medskip

If \texttt{UseSpecialAlgorithms} is set to \texttt{true} and $f$ does not fall into any of the aforementioned cases, then the function \texttt{fpt} next calls \texttt{isSimpleNormalCrossing(f)} (see Section~\ref{subsec.SNC}) to check whether the polynomial $f$ is (locally, at the origin) a product of factors that are in simple normal crossing, in which case the $F$-pure threshold is simply the reciprocal of the largest multiplicity occurring in that factorization.  
% If \texttt{isSimpleNormalCrossing(f)} outputs true, then the $F$-pure threshold is easily computed, since is just the minimum of the values $1/a_i$, for $1 \leq i \leq n$.
Note that the function \texttt{factor} is called whenever \texttt{isSimpleNormalCrossing} is used, which can sometimes make the verification slow.  The user can avoid this by setting \texttt{UseSpecialAlgorithms} to \texttt{false}.

@@

In the remainder of this section, we explain how the function  \texttt{fpt} proceeds when no special algorithm is available, or when \texttt{UseSpecialAlgorithms} is set to \texttt{false}, and describe the roles of the options \texttt{DepthOfSearch} and \texttt{Attempts}, whose values are both nonnegative integers.

In this case, the \texttt{fpt} function either computes, and outputs, the exact value of the $F$-pure threshold of $f$, or otherwise recursively computes a finite sequence of lower and upper bounds for this value, and outputs the last of these, which will be the tightest among all computed.  The value of the option \texttt{DepthOfSearch} determines the precision of the initial set of bounds, and the option \texttt{Attempts}, in conjunction with some various subroutines, determines whether, and how, to produce new, tighter bounds from the previous ones.

More specifically, let $e$ denote the value of the option \texttt{DepthOfSearch}, which conservatively defaults to \texttt{1}.  The \texttt{fpt} function first computes $\nu=\nu_f(p^e)$, which agrees with the output of \texttt{nu(e,f)}.  It is well known that the $F$-pure threshold of $f$ is greater than $\nu/p^e$ and at most $(\nu+1)/p^e$, and applying  \cite[Proposition 4.2]{HernandezFPurityOfHypersurfaces} to this lower bound tells us that the $F$-pure threshold of $f$ must be at least $\nu/(p^e-1)$.  In summary, we know that the $F$-pure threshold of $f$ must lie in the closed interval
%
\begin{equation}
\label{estimating-interval: e}
\tag{$\dagger$}
\left[ \frac{\nu}{p^e-1}, \frac{\nu+1}{p^e} \right].
\end{equation}

@@

\medskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i9 : f = x^2*(x + y)^3*(x + 3*y^2)^5;

@@

i12 : fpt(f, Attempts => 1, DepthOfSearch => 3) -- the right-hand
      endpoint (nu+1)/p^e is the F-pure threshold

@@

o12 : QQ
\end{MyVerbatim}
}
\medskip

@@

\medskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i13 : f = x^6*y^4 + x^4*y^9 + (x^2 + y^3)^3;

@@

i15 : fpt(f, Attempts => 2, DepthOfSearch => 3) -- the left-hand endpoint
      nu/(p^e-1) is the F-pure threshold

@@

o15 : QQ
\end{MyVerbatim}
}
\medskip

@@

\medskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i12 : f = x^3*y^11*(x + y)^8*(x^2 + y^3)^8;

i13 : fpt(f, DepthOfSearch => 3, Attempts => 4)

        1   4
o13 = {--, --}
       20  75

o13 : List

i14 : fpt(f, DepthOfSearch => 3, Attempts => 6)

        13   4
o14 = {---, --}
       250  75

i15 : fpt(f, DepthOfSearch => 3, Attempts => 8)

       1
o15 = --
      19

o15 : QQ
\end{MyVerbatim}
}
\medskip

@@

\medskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i20 : f = x^7*y^5*(x + y)^5*(x^2 + y^3)^4;

@@

        3   1
o21 = {--, --}
       40  13

@@

        24   1
o22 = {---, --}
       313  13

o22 : List
\end{MyVerbatim}
}
\medskip

% f = x^7*y^5*(x + y)^5*(x^2 + y^3)^4;
% fpt(f, DepthOfSearch => 3, Attempts => 5)
% fpt(f, DepthOfSearch => 3, Attempts => 5, Bounds => oo)

If \texttt{guessFPT} is unsuccessful and \texttt{FinalAttempt} is set to \texttt{true}, the \texttt{fpt} function proceeds to use the convexity of the $F$-signature function and a secant line argument to attempt to narrow down the interval bounding the $F$-pure threshold.
If successful, the new lower bound may coincide with the upper bound, in which case we can conclude that it is the desired $F$-pure threshold.
If this is not the case, an $F$-regularity check is done at the new lower bound, to verify if it is the $F$-pure threshold.

\medskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i2 : f = 2*x^10*y^8+x^4*y^7-2*x^3*y^8;

i3 : numeric fpt(f, DepthOfSearch => 3)

o3 = {.14, .144}

o3 : List

i4 : numeric fpt(f, DepthOfSearch => 3, FinalAttempt => true)
     -- FinalAttempt improves the estimate slightly

o4 = {.142067, .144}

o4 : List
\end{MyVerbatim}
}
\medskip

The computations performed when \texttt{FinalAttempt} is set to \texttt{true} are often slow, and often fail to improve the estimate, and for this reason, this option should be used sparingly. It is often more effective to increase the values of \texttt{Attempts} or \texttt{DepthOfSearch}, instead.

\medskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i35 : time numeric fpt(f, DepthOfSearch => 3, FinalAttempt => true)
     -- used 0.684759 seconds

o35 = {.142067, .144}

o35 : List

i36 : time fpt(f, DepthOfSearch => 3, Attempts => 7) 
     -- used 0.405148 seconds

      1
o36 = -
      7

o36 : QQ

i37 : time fpt(f, DepthOfSearch => 4) 
     -- used 0.365025 seconds

      1
o37 = -
      7

o37 : QQ
\end{MyVerbatim}
}
\medskip

@@

\medskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i39 : fpt(f, DepthOfSearch => 3, FinalAttempt => true, Verbose => true)

@@

        8009   18
o39 = {-----, ---}
       56375  125

o39 : List
\end{MyVerbatim}
}
\medskip

@@

\medskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i31 : R = ZZ/7[x,y];

i32 : f = x*(y - 1)^2 - y*(x - 1)^3;

@@

o33 : QQ

i34 : fpt(f)

o34 = 1
\end{MyVerbatim}
}
\medskip

\noindent In this case, most options enabled by \texttt{UseSpecialAlgorithms => true}
are ignored except for the check for simple normal crossings.  \texttt{FinalCheck => true}
is also ignored.

\subsection{The function \texttt{isSimpleNormalCrossings}} \label{subsec.SNC}
As mentioned earlier, the function \texttt{isSimpleNormalCrossings} verifies whether $f$ is a product of factors in simple normal crossing.
Suppose that $f$ has factorization $f_1^{a_i} f_2^{a_2} \cdots f_n^{a_n}$.  Recall that its factors $f_i$ are said to be in
\emph{simple normal crossing} at a point if, locally, they form part of a regular system of parameters.  The function \texttt{isSimpleNormalCrossings} determines whether $f$ has simple normal crossings by computing the Jacobian matrix of each subset of $\{ f_i \}_{i=1}^n$ (evaluated at the origin), and checking that these matrices have the expected rank, and that these subsets generate ideals of the appropriate height.

\medskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i36 : R = ZZ/7[x, y, z];

i37 : isSimpleNormalCrossing(x^2 - y^2)

o37 = true

i38 : isSimpleNormalCrossing(x^2 - y*z)

o38 = false
\end{MyVerbatim}
}
\medskip

>>>>

Consider a nonzero polynomial $f$ and a natural number $e$.  If $f$ does not vanish at the origin, then set $\nu_f^{\idealm}(p^e) := \infty$.  Otherwise, $f \in \idealm$, and we instead define
\[ \nu_f^{\idealm}(p^e) \coloneqq  \max \{ n \in \mathbb{N} : f^n \notin \idealm^{[p^e]} \}, \]
where $\idealm^{[p^e]}$ denotes the $p^e$-th Frobenius power of $\idealm$, that is, the ideal generated by the $p^e$-th powers of the elements of $\idealm$.
%It is not hard to see that this number is the largest power of $f$ that is part of a free basis for the localization of the ambient polynomial ring at the origin, over its subring of $p^e$-th powers.
%As noted earlier, this number also has a natural interpretation in terms of $F$-purity of pairs.
Following our outline, we define
\[ \ft^{\idealm}(f) \coloneqq  \lim_{e \to \infty} \frac{ \nu_f^{\idealm}(p^e)}{p^e}. \]

@@

We conclude this subsection by briefly reviewing some natural generalizations.
Suppose that $I$ and $J$ are ideals.
If $I$ is contained in the radical of $J$, then we set
%
\[ \nu_I^J(p^e) \coloneqq \max \{ n \in \mathbb{N}: I^n \not\subseteq J^{[p^e]} \}, \]
or $\nu_I^J(p^e) \coloneqq 0$, when the set on the right-hand side is empty.
Otherwise, we set $\nu_I^J(p^e) \coloneqq \infty$.
This clearly generalizes the quantity $\nu_f^{\idealm}(p^e)$ considered earlier, and we call
\[ \ft^J(I) \coloneqq  \lim_{e \to \infty} \frac{ \nu_I^J(p^e)}{p^e} \]
the \emph{$F$-threshold of $I$ with respect to $J$}. 
The value $\ft^{\idealm}(I)$ is called the \emph{$F$-pure threshold of $I$} at the origin, and if $I = \langle f \rangle$ is principal,  $\ft^{J}(f) \coloneqq  \ft^J(I)$ is called the \emph{$F$-threshold of $f$ with respect to $J$}.
Like $F$-pure thresholds, $F$-thresholds are rational (when finite), and can be characterized in terms of test ideals.

@@

\smallskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i1 : R = ZZ/11[x,y];

@@

i4 : nu(2, I, J)

o4 = 281

@@

i6 : nu(2, f, J)

o6 = 120
\end{MyVerbatim}
}

@@

We now describe the options available for the function \texttt{nu}.
As pointed out in the introduction, if $f \in \mathfrak{m}$, the values $\nu^{\mathfrak{m}}_f(p^e)$ can be recovered from the $F$-pure threshold of $f$.
This is used to speed up computations for certain polynomials whose $F$-pure thresholds can be computed quickly via specialized algorithms or formulas, namely diagonal polynomials, binomials, forms in two variables, and products of factors in simple normal crossing (see Section~\ref{sec.FPT}).
This feature can be disabled by setting the option \texttt{UseSpecialAlgorithms} (default value \texttt{true}) to \texttt{false}.\footnote{In Section~\ref{ss: UseSpecialAlgorithms} we discuss a couple of situations in which this may be desirable.}

\smallskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i7 : R = ZZ/17[x,y,z];

i8 : f = x^3 + y^4 + z^5; -- a diagonal polynomial

i9 : time nu(10, f)
     -- used 0.0161622 seconds

o9 = 1541642394460

i10 : time nu(10, f, UseSpecialAlgorithms => false)
     -- used 2.06877 seconds

o10 = 1541642394460
\end{MyVerbatim}
}
\smallskip

@@

\smallskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i11 : R = ZZ/5[x,y,z];

i12 : m = ideal(x, y, z);

o12 : Ideal of R

i13 : time nu(2, m, m^2) -- uses binary search (default)
     -- used 1.82479 seconds

o13 = 97

i14 : time nu(2, m, m^2, Search => Linear) -- but linear seach gets luckier
     -- used 0.597035 seconds

o14 = 97
\end{MyVerbatim}
}
%\medskip
\smallskip

@@

% \medskip
\smallskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i15 : nu(5, x^2*y^4 + y^2*z^7 + z^2*x^8, ReturnList => true)

o15 = {0, 1, 8, 44, 224, 1124}

o15 : List
\end{MyVerbatim}
}
%\medskip
\smallskip

@@

% \medskip
\smallskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i16 : R = ZZ/11[x,y,z];

i17 : f = x^3 + y^3 + z^3 + x*y*z;

i18 : time nu(3, f) -- ContainmentTest is set to FrobeniusRoot, by default
     -- used 0.153691 seconds

o18 = 1209

i19 : time nu(3, f, ContainmentTest => StandardPower)
     -- used 10.1343 seconds

o19 = 1209
\end{MyVerbatim}
}
%\medskip

@@

\smallskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i20 : R = ZZ/3[x,y];

i21 : m = ideal(x, y);

i22 : nu(4, m^5)

o22 = 32

i23 : nu(4, m^5, ContainmentTest => FrobeniusPower)

o23 = 26
\end{MyVerbatim}
}
\smallskip

@@

\smallskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i24 : R = ZZ/7[x,y];

i25 : f = (x - 1)^3 - (y - 2)^2;

i26 : nu(3, f)

o26 = infinity

o26 : InfiniteNumber

i27 : nu(3, f, AtOrigin => false)

o27 = 285
\end{MyVerbatim}
}

@@

The \emph{FrobeniusThresholds} package contains methods to test candidate values for $F$-pure thresholds and $F$-jumping numbers, even in some singular rings.
Consider a  $\mathbb{Q}$-Gorenstein ring $R$ of characteristic $p>0$, whose index is not divisible by $p$.
Given a parameter $t\in \mathbb{Q}$ and an element $f$ of $R$, the command \texttt{isFPT(t, f)} checks whether $t$ is the $F$-pure threshold of $f$, while \texttt{compareFPT(t, f)} provides further information, returning {\tt-1}, \texttt{0}, or \texttt{1} when $t$ is, respectively, less than, equal to, or greater than the $F$-pure threshold of $f$.  Setting the option \texttt{AtOrigin} to \texttt{true} tells the function to consider the $F$-pure threshold at the origin.  % The following example illustrates this functionality.

\smallskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i1 : R = ZZ/11[x,y,z]/(x^2 - y*(z - 1));

@@

o4 = false
\end{MyVerbatim}
}
\smallskip

The general method applied calls upon functionality from the \emph{TestIdeals} package.
The functions \texttt{testIdeal} and \texttt{FPureModule} therein are used to compute the test ideals $\tau(f^t)$ and $\tau(f^{t-\varepsilon})$, for $0<\varepsilon\ll 1$.
By comparing these test ideals, we can decide whether $t$ is the $F$-pure threshold of $f$.
For instance, when $R$ is a polynomial ring, $t$ is the $F$-pure threshold of $f$ at the origin if and only if $\tau(f^t)$ is contained in the homogeneous maximal ideal, but $\tau(f^{t-\varepsilon})$ is not.

Since not only the $F$-pure thresholds, but also the higher $F$-jumping numbers, are determined by containment conditions on test ideals, the functionality is extended to determine whether a given number is an $F$-jumping number.

\smallskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i5 : R = ZZ/13[x,y];

@@

o7 = true

i8 : isFPT(3/4, f)

o8 = false
\end{MyVerbatim}
}

@@

\smallskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i1 : R = ZZ/5[x,y,z];

@@

The \texttt{fpt} function has an option called \texttt{UseSpecialAlgorithms}, which, when set to \texttt{true} (its default value), tells \texttt{fpt} to first check whether $f$ is a diagonal polynomial, a binomial, a form in two variables, or a product of factors in simple normal crossing, in that order.
When $f$ is a diagonal polynomial, a binomial, or a form in two variables, algorithms of Hern\'andez \cite{HernandezFInvariantsOfDiagonalHyp, HernandezFPureThresholdOfBinomial} or Hern\'andez and Teixeira \cite{HernandezTeixeiraFThresholdFunctions} are executed to compute the $F$-pure threshold.
%Algorithms for computing $F$-pure thresholds of binomials also appear in \cite{ShibutaTakagiLCThresholds}, though they are only valid in certain special characteristics.

\smallskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i4 : fpt(x^17 + y^20 + z^24) -- a diagonal polynomial

@@

o7 : QQ
\end{MyVerbatim}
}
\smallskip

\noindent The algorithm for computing the $F$-pure threshold of a binary form $f$ requires factoring $f$ into linear forms, and that may be difficult or impossible when that factorization occurs in a Galois field of excessively large order.
This is a situation when the user will want to set the option \texttt{UseSpecialAlgorithms} to \texttt{false}.
However, when a factorization is already known, instead of passing $f$ to \texttt{fpt}, the user can pass a list containing all the pairwise coprime linear factors of $f$ to \texttt{fpt}, and a list containing their respective multiplicities.

\smallskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i8 : fpt({x, y, x + y, x + 3*y}, {2, 6, 9, 10}) == oo

o8 = true
\end{MyVerbatim}
}
\smallskip

If \texttt{UseSpecialAlgorithms} is set to \texttt{true} and $f$ does not fall into any of the aforementioned cases, then the function \texttt{fpt} next calls \texttt{isSimpleNormalCrossing(f)} (see Section~\ref{subsec.SNC}) to check whether the polynomial $f$ is (locally, at the origin) a product of factors that are in simple normal crossing, in which case the $F$-pure threshold is simply the reciprocal of the largest multiplicity occurring in that factorization.  
% If \texttt{isSimpleNormalCrossing(f)} outputs true, then the $F$-pure threshold is easily computed, since is just the minimum of the values $1/a_i$, for $1 \leq i \leq n$.
Note that the function \texttt{factor} is called whenever \texttt{isSimpleNormalCrossing} is used, and that can sometimes make the verification slow.  The user can avoid this by setting \texttt{UseSpecialAlgorithms} to \texttt{false}.

@@

We now explain how the function  \texttt{fpt} proceeds when no special algorithm is available, or when \texttt{UseSpecialAlgorithms} is set to \texttt{false}.
In this case, \texttt{fpt} computes a sequence of lower and upper bounds for the $F$-pure threshold of $f$, and either finds its exact value in this process, or outputs the last of these sets of bounds, which will be the tightest among all computed.
The value of the option \texttt{DepthOfSearch} determines the precision of the initial set of bounds, and the option \texttt{Attempts} determines, roughly, how many new, tighter sets of bounds are to be computed.

More specifically, let $e$ denote the value of the option \texttt{DepthOfSearch}, which conservatively defaults to \texttt{1}.
The \texttt{fpt} function first computes $\nu=\nu_f(p^e)$, which agrees with the output of \texttt{nu(e,f)}.
It is well known that the $F$-pure threshold of $f$ is greater than $\nu/p^e$ and at most $(\nu+1)/p^e$, and applying  \cite[Proposition 4.2]{HernandezFPurityOfHypersurfaces} to the lower bound tells us that the $F$-pure threshold of $f$ must be at least $\nu/(p^e-1)$.
In summary, we know that the $F$-pure threshold of $f$ must lie in the closed interval
%
\begin{equation}
\label{estimating-interval: e}
\tag{$\dagger$}
\left[ \frac{\nu}{p^e-1}, \frac{\nu+1}{p^e} \right].
\end{equation}

@@

\smallskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i9 : f = x^2*(x + y)^3*(x + 3*y^2)^5;

@@

i12 : fpt(f, Attempts => 1, DepthOfSearch => 3) -- the right-hand endpoint
      (ν+1)/p^e is the F-pure threshold

@@

o12 : QQ
\end{MyVerbatim}
}
\smallskip

@@

\smallskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i13 : f = x^6*y^4 + x^4*y^9 + (x^2 + y^3)^3;

@@

i15 : fpt(f, Attempts => 2, DepthOfSearch => 3) -- the left-hand endpoint
      ν/(p^e-1) is the F-pure threshold

@@

o15 : QQ
\end{MyVerbatim}
}
\smallskip

@@

\smallskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i16 : f = x^3*y^11*(x + y)^8*(x^2 + y^3)^8;

i17 : fpt(f, DepthOfSearch => 3, Attempts => 4)

        1   4
o17 = {--, --}
       20  75

o17 : List

i18 : fpt(f, DepthOfSearch => 3, Attempts => 6)

        13   4
o18 = {---, --}
       250  75

o18 : List

i19 : fpt(f, DepthOfSearch => 3, Attempts => 8)

       1
o19 = --
      19

o19 : QQ
\end{MyVerbatim}
}
\smallskip

@@

\smallskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i20 : f = x^7*y^5*(x + y)^5*(x^2 + y^3)^4;

@@

        19   1
o21 = {---, --}
       250  13

@@

        45   1
o22 = {---, --}
       589  13

o22 : List
\end{MyVerbatim}
}
\smallskip

If \texttt{guessFPT} is unsuccessful and \texttt{FinalAttempt} is set to \texttt{true}, the \texttt{fpt} function proceeds to use the convexity of the $F$-signature function and a secant line argument to attempt to narrow down the interval bounding the $F$-pure threshold.
If successful, the new lower bound may coincide with the upper bound, in which case we can conclude that it is the desired $F$-pure threshold.
If this is not the case, a check is performed to verify if the new lower bound is the $F$-pure threshold.

\smallskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i23 : f = 2*x^10*y^8 + x^4*y^7 - 2*x^3*y^8;

i24 : numeric fpt(f, DepthOfSearch => 3)

o24 = {.14, .144}

o24 : List

i25 : numeric fpt(f, DepthOfSearch => 3, FinalAttempt => true)
      -- FinalAttempt improves the estimate slightly

o25 = {.142067, .144}

o25 : List
\end{MyVerbatim}
}
\smallskip

The computations performed when \texttt{FinalAttempt} is set to \texttt{true} are often slow, and often fail to improve the estimate, and for this reason, this option should be used sparingly.
It is typically more effective to increase the values of \texttt{Attempts} or \texttt{DepthOfSearch}, instead.

\smallskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i26 : time numeric fpt(f, DepthOfSearch => 3, FinalAttempt => true)
     -- used 0.72874 seconds

o26 = {.142067, .144}

o26 : List

i27 : time fpt(f, DepthOfSearch => 3, Attempts => 7) 
     -- used 0.452872 seconds

      1
o27 = -
      7

o27 : QQ

i28 : time fpt(f, DepthOfSearch => 4) 
     -- used 0.338834 seconds

      1
o28 = -
      7

o28 : QQ
\end{MyVerbatim}
}
\smallskip

@@

\smallskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i29 : fpt(f, DepthOfSearch => 3, FinalAttempt => true, Verbose => true)

@@

        8009   18
o29 = {-----, ---}
       56375  125

o29 : List
\end{MyVerbatim}
}
\smallskip

@@

\smallskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i30 : R = ZZ/7[x,y];

i31 : f = x*(y - 1)^2 - y*(x - 1)^3;

i32 : fpt(f)

o32 = 1

@@

o33 : QQ
\end{MyVerbatim}
}
\smallskip

\noindent In this case, most features enabled by \texttt{UseSpecialAlgorithms => true}
are ignored, except for the check for simple normal crossings;  \texttt{FinalCheck => true}
is also ignored.

\subsection{The function \texttt{isSimpleNormalCrossings}} \label{subsec.SNC}
As mentioned earlier, the function \texttt{isSimpleNormalCrossings} verifies whether a ring element $f$ is a product of factors in simple normal crossing.
Suppose that $f$ has factorization $f_1^{a_i} f_2^{a_2} \cdots f_n^{a_n}$.  Recall that its factors $f_i$ are said to be in
\emph{simple normal crossing} at a point if, locally, they form part of a regular system of parameters.  The function \texttt{isSimpleNormalCrossings} determines whether $f$ has simple normal crossings at the origin by computing the Jacobian matrix of each subset of $\{ f_1, \ldots, f_n \}$ (evaluated at the origin), and checking that these matrices have the expected rank, and that these subsets generate ideals of the appropriate height.

\smallskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i34 : R = ZZ/7[x,y,z];

i35 : isSimpleNormalCrossing(x^2 - y^2)

o35 = true

i36 : isSimpleNormalCrossing(x^2 - y*z)

o36 = false
\end{MyVerbatim}
}
\smallskip

The function \texttt{isSimpleNormalCrossing} is exposed to the user, so can be used independent of any $F$-pure threshold calculation.
If the user sets its option \texttt{AtOrigin} to \texttt{false} (its default value is \texttt{true}), then the function checks whether the $f_i$ are in simple normal crossing \emph{everywhere}, which can be much slower, since Jacobian ideals are computed.

\smallskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i37 : R = QQ[x,y,z];
